\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2017-01-20 mSystems/rev/mSys_MandM_rev.tex
Corrections and pre-submission, which leads to REVISION draft I
Changes to be committed:
	modified:   figs/supfig_taxlev1.eps
	modified:   figs/supfig_taxlev2.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	new file:   tab/mSys_suppl_tab1.pdf
	new file:   tab/mSys_suppl_tab2.pdf
	new file:   tab/mSys_suppl_tab3.pdf
	new file:   tab/mSys_suppl_tab4.pdf
	new file:   tab/mSys_suppl_tab5.pdf
	new file:   tab/m

%%
\subsection*{Standardization} \label{sec:stan}
In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for each study independently. With this approach, all the studies can be visualized in a shared plot with units of Taylor--parameter standard--deviation on their axes.

@@

The bacteria and archaea taxonomic assignations were obtained by analyzing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97\% of their sequence identity using QIIME \cite{QIIME}. Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analyzed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold. Genus, with the best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We verified that our conclusions were not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure \ref{supfig:taxlev1}).

%%
\subsubsection*{Sample selection}
We chose studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states. Only those individuals who had three or more time points of data available in databases were selected. The study by Caporaso \textit{et al.} study \cite{moving} was selected as it featured two healthy individuals measured over a very long timespan, with almost daily sampling. The study of Faith \textit{et al.} \cite{LEA} was selected given the BMI differences between subjects. Moreover, some of them followed diets which could be treated as system perturbations. Only those individuals who had normal or overweight BMI were considered as healthy. The study by Smith \textit{et al.} \cite{kwashiorkor} was selected for both the age of the patients and the rare disease. We only worked with the discordant twins, and considered those who were not affected by kwashiorkor in each pair of patients as being healthy. The study by David \textit{et al.} \cite{diet}was selected for its differential diets. The healthy part was considered to be the first time samples of each individual before the diet, while the rest of the time points were considered as perturbations. Dethlefsen and Relman's work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes by three different subjects of the same antibiotic. The healthy part was considered to be only those times before any antibiotic treatment, and the time of antibiotic intakes and the period after that as perturbations. The work by David \textit{et al.} \cite{hostlife} was selected due to the comprehensive longitudinal data that it provides plus its complete metadata and the interesting events that happened to both subjects (an infection and a trip abroad). The healthy part was taken as the previous time points before each event. Finally, we also considered a study from our group carried out by Durban \textit{et al.} \cite{IBS} in which the healthy subjects were considered as those who did not suffer from irritable bowel syndrome, while the patients who had this disease were taken as perturbations.

The metadata for each study is provided in Supplementary Tables S1 to S7. They all used 16S rRNA gene sequencing, except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, in which both SMS and 16S rRNA data were used. In the latter case, we chose to work with SMS data to show that our method was valid, regardless of the source of taxonomic information. Each of the datasets was treated as follows:

@@

%%
\subsubsection*{Taxa level robustness}
We selected genus as the taxonomic level for the subsequent steps of our work. In order to ensure that there were no crucial differences between adjacent taxonomic levels which could still be of relevance after standardization (see the last subsection of Material and Methods), we tested two different data sets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Supplementary Figures \ref{supfig:taxlev1} (overview) and \ref{supfig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

%TC:endignore
\begin{supfig} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
\caption{Overview of the comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. The former row of subfigures is for 16S, where levels are family (blue circles) vs. genus (purple triangles), whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs. species (purple triangles). The left column shows the raw results and the right column plots the standardized results (see Standardization in Material and Methods).}
\label{supfig:taxlev1}
\end{supfig}

\begin{supfig} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
  \includegraphics[width=0.99\textwidth]{figs/supfig_taxlev1.eps}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{supfig:taxlev2}
\end{supfig}
%TC:ignore

@@

The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} and Supplementary Figures \ref{supfig:corrank_HLS_before} and \ref{supfig:corrank_HLS_after}. The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes over time. So, the RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
\begin{linenomath}
$${\rm RSI} = \left(1-\frac{\text{true rank hops}}{\text{possible rank hops}}\right)^p = \left(1-\frac{D}{(N-1)(t-1)}\right)^p$$
\end{linenomath}
where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples. The power index $p=4$ was arbitrarily chosen to increase the resolution in the stable region.

>>>>

%%
\subsection*{Standardization} \label{sec:stan}
In order to properly show all the studies under common axes, we decided to standardize the Taylor parameters using the group of healthy individuals for every single study independently. With this approach, all the studies can be visualized in a shared plot with units of Taylor--parameter standard--deviation on their axes.

@@

The bacteria and archaea taxonomic assignations were obtained by analyzing 16S rRNA sequences, which were clustered into operational taxonomic units (OTUs) sharing 97\% of their sequence identity using QIIME \cite{QIIME}. Shotgun metagenomic sequencing (SMS) data\cite{kwashiorkor} were analyzed and assigned at strain level by the Livermore Metagenomic Analysis Toolkit (LMAT)\cite{LMAT}, according to their default quality threshold. Genus, with the best balance between error assignment and number of taxa, was chosen as our reference taxonomic level. We verified that our conclusions were not significantly affected by selecting family or species as the reference taxonomic level (see Supplementary Figure S\ref{supfig:taxlev1}).

%%
\subsubsection*{Sample selection}
We chose studies about relevant pathologies containing metagenomic sequencing time data series of bacterial populations from humans in different healthy and non-healthy states. Only those individuals who had three or more time points of data available in databases were selected. The study by Caporaso \textit{et al.} study \cite{moving} was selected as it featured two healthy individuals measured over a very long timespan, with almost daily sampling. The study of Faith \textit{et al.} \cite{LEA} was selected given the BMI differences between subjects. Moreover, some of them followed diets which could be treated as system perturbations. Only those individuals who had normal or overweight BMI were considered as healthy. The study by Smith \textit{et al.} \cite{kwashiorkor} was selected for both the age of the patients and the rare disease. We only worked with the discordant twins, and considered those who were not affected by kwashiorkor in each pair of patients as being healthy. The study by David \textit{et al.} \cite{diet} was selected for its differential diets. The healthy part was considered to be the first time samples of each individual before the diet, while the rest of the time points were considered as perturbations. Dethlefsen and Relman's work \cite{antibiotic} was selected due to the interesting treatment of two antibiotic intakes by three different subjects of the same antibiotic. The healthy part was considered to be only those times before any antibiotic treatment, and the time of antibiotic intakes and the period after that as perturbations. The work by David \textit{et al.} \cite{hostlife} was selected due to the comprehensive longitudinal data that it provides plus its complete metadata and the interesting events that happened to both subjects (an infection and a trip abroad). The healthy part was taken from time points before or after each event. Finally, we also considered a study from our group carried out by Durban \textit{et al.} \cite{IBS} in which the healthy subjects were considered as those who did not suffer from irritable bowel syndrome, while the patients who had this disease were taken as perturbations.

The metadata for each study is provided in Supplementary Tables S\ref{tab:diet} to S\ref{tab:LEA}. They all used 16S rRNA gene sequencing, except for the study of the discordant kwashiorkor twins\cite{kwashiorkor}, in which both SMS and 16S rRNA data were used. In the latter case, we chose to work with SMS data to show that our method was valid, regardless of the source of taxonomic information. Each of the datasets was treated as follows:

@@

%%
\subsubsection*{Taxa level robustness}
We selected genus as the taxonomic level for the subsequent steps of our work. In order to ensure that there were no crucial differences between adjacent taxonomic levels which could still be of relevance after standardization (see the last subsection of Material and Methods), we tested two different data sets. In the former, the antibiotics study\cite{antibiotic} with 16S data, we tested the differences between genus and family levels. The latter dataset tested was the kwashiorkor discordant twins study\cite{kwashiorkor} for both genus and species taxonomic levels. The Supplementary Figures S\ref{supfig:taxlev1} (overview) and S\ref{supfig:taxlev2} (detail) plot the comparison between studies (and so, 16S and SMS) and between adjacent taxonomic levels.

%TC:endignore
\begin{supfig} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_raw_SMS.pdf} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/sum_sta_SMS.pdf}
  \includegraphics[width=1.0\textwidth]{figs/supfig_taxlev1.eps}
\caption{Overview of the comparison of different approaches based on adjacent taxonomic levels using plots in the Taylor-parameters space. The former row of subfigures is for 16S, where levels are family (blue circles) vs. genus (purple triangles), whereas the latter row of subfigures is for SMS, where levels are genus (blue circles) vs. species (purple triangles). The left column shows the raw results and the right column plots the standardized results (see Standardization in Material and Methods).}
\label{supfig:taxlev1}
\end{supfig}

\begin{supfig} 
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_fam_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_16S.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_gen_SMS.pdf}
  %\includegraphics[width=0.5\textwidth]{results/taxalevel/xWb_spc_SMS.pdf}
  \includegraphics[width=1.0\textwidth]{figs/supfig_taxlev2.eps}
\caption{Detail of comparison of different approaches based on adjacent taxonomic levels using plots of X-weighted power-law fits (see Material and Methods). The former row of subfigures shows examples for 16S, whereas the latter row of subfigures plots examples for SMS. The left column shows results for the superior taxonomic level (family for 16S, genus for SMS), while the right column shows results for the inferior level (genus for 16S, specie for SMS).}
\label{supfig:taxlev2}
\end{supfig}
%TC:ignore

@@

The Rank Stability Index (RSI) is shown as a percentage in a separate bar on the right of the rank matrix plot in Figures \ref{fig:corrank_HLS_abroad} and \ref{fig:corrank_HLS_returned} and Supplementary Figures S\ref{supfig:corrank_HLS_before} and S\ref{supfig:corrank_HLS_after}. The RSI is strictly $1$ for an element whose range never changes over time, and is strictly $0$ for an element whose rank oscillates between the extremes over time. So, the RSI is calculated, per element, as $1$ less the quotient of the number of true rank hops taken between the number of maximum possible rank hops, all powered to $p$:
\begin{linenomath}
$${\rm RSI} = \left(1-\frac{\text{true rank hops}}{\text{possible rank hops}}\right)^p = \left(1-\frac{D}{(N-1)(t-1)}\right)^p$$
\end{linenomath}
where $D$ is the total of rank hops taken by the studied element, $N$ is the number of elements that have been ranked, and $t$ is the number of time samples. The power index $p=4$ was arbitrarily chosen to increase the resolution in the stable region.

Finally, under the rank matrix of the aforementioned Figures there are plots relevant to the variability of the rank over time. On the one hand, the RV (Rank Variability) for a sampling point shows the absolute difference between every taxon?s rank and its accumulated abundance rank (the overall rank), averaged for all the taxa shown. On the other hand, the DV (Differences Variability) for a sampling point shows the absolute difference between every taxon?s rank at that time and the value that it had at the previous sampling point, averaged for all the taxa shown.

### 2017-01-20 mSystems/rev/mSys_results_rev.tex
mSys: Sync of Word and LaTeX versions after sending REV1
Changes to be committed:
	modified:   mSys_MAIN_rev.docx
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_intro_rev.tex
	modified:   mSys_results_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex

The microbiome temporal variability was analyzed to extract the global properties of the system. As fluctuations in total counts are plagued by systematic errors we worked on the temporal variability of relative abundances for each taxon. Our first finding was that, in all cases, changes in the relative abundances of taxa followed a ubiquitous pattern, known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., the microbiota of all detected taxa followed $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between the mean relative abundance $x_i$ and the dispersion $\sigma_i$. The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances. As shown in Figure \ref{fig:main1}, the most abundant species were less volatile in relative terms than those which were less abundant. The fitting to the power law was always robust ($R^{2}$ > 0.88) and did not depend on the microbiome condition. The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with the stability of the community. The variability V was a direct estimator of the amplitude of fluctuations over time, while the scaling index gave us information about the statistical properties of the ecosystem. The fact that it was less than the unity (when looking at the standard deviation) tells us that the most abundant taxa were less susceptible to any perturbation than the less abundant elements of the population. The Taylor parameters were related to the health status of the host, which we consider as constituting the main finding contributed by this article.

\begin{figure}
	\centering
	\vspace*{-15mm} % Corrects overbox of the figures
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{figs/Fig1.eps}
	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored over time. The fit is shown for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}. Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude.}
	\label{fig:main1}
\end{figure}

Taylor's parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S1 to S7. Our results hint at ubiquitous behavior. Firtly, the variability (which corresponds to the maximum amplitude of fluctuations) was large, which suggests the resilient capacity of the microbiota. Secondly, the scaling index was always smaller than one, which means that more abundant taxa were less volatile than less abundant ones. In addition, Taylor's parameters for the microbiome of healthy individuals in different studies were compatible with estimated errors. This enabled us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.

In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for each study independently (for details of the procedure, please see Standardization subsection in Material and Methods). The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota was compromised (i.e., they were suffering from IBS, kwashiorkor, altered diet, intake of antibiotics, a Salmonella infection, or had gone on a trip abroad) are shown in Figure \ref{fig:main2}. The variability in children developing kwashiorkor was smaller than that of their healthy twins. A meat/fish-based diet significantly increased variability when compared to a plant-based diet. All other cases presented increased variability, which was particularly severe and statistically significant at over 95\% confidence level (CL), for grade III obese patients on a diet, individuals taking antibiotics, the subject who had a Salmonella infection, the subject who did a travel abroad or the IBS-diagnosed patients. One global property emerged from all the worldwide data collected: the Taylor's parameters characterized the statistical behavior of microbiome changes. Furthermore, we verified that our conclusions were robust to systematic errors as a result of taxonomic assignment (see Taxa level selection in Material and Methods).

\begin{figure}
	\centering
	%\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
	\includegraphics[width=1.0\textwidth]{figs/Fig2.eps}
	\caption{Taylor's law parameter space. All the data studied in this work were compiled here. The colored circle corresponds to a 68\% confidence level (CL) region of healthy individuals in the Taylor's parameter space, while the dashed line delimites the 98\% CL region. Points with errors place gut microbiome in the Taylor's parameter space, for each individual whose microbiota was compromised. It should be noted that the parameters were standardized (standard deviation units) to the healthy group in each study for demonstrative and comparative purposes.}
	\label{fig:main2}
\end{figure}

Taylor's power law has been explained in terms of various effects, though none have brought a general consensus. It can be shown to have its origin in mathematical convergence which is similar to the central limit theorem, and thus virtually any statistical model designed to produce a Taylor law converges to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}. To reveal the generic mechanisms that drive different scenarios in the $\beta-V$ space, we modeled the system by assuming that taxon relative abundance followed a Langevin equation with, on the one hand, a deterministic term that captured the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}. Both terms were modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods). In this model, when $V$ is sufficiently low, abundances are stable in time. Differences in the variability $V$ can induce a noise-induced phase transition in the relative abundances of taxa. The temporal evolution of the probability of a taxon having the abundance $x_i$, given its fitness, is governed by the Fokker--Planck equation. The results of solving this equation show that stability is best captured by a phase space determined by the fitness $F$ and the amplitude of fluctuations $V$ (see Figure \ref{fig:main3}).

@@

The model predicted two phases for the gut microbiome: a stable phase with large variability that enabled some changes in the relative abundances of taxa; and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly over time. The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase. In particular, individuals taking antibiotics and the IBS--diagnosed patient P2 had the most severe symptoms. In this phase diagram, each microbiota state is represented by a point at its measured variability $V$ and inferred fitness $F$. The model predicted high average fitness for all taxa, i.e., taxa were narrowly distributed in F. The fitness parameter was chosen with different values for demonstrative purposes. Fitness was larger for the healthiest subjects and smaller for the IBS--diagnosed patients.

@@

We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure \ref{supfig:corrank_HLS_before} and Supplementary Figure \ref{supfig:corrank_HLS_after} for the corresponding rank plots). See Supplementary Table S8 for details about the rank and RSI for the above-mentioned taxa over the different periods considered.

@@

Finally, we studied the time dependence of the variability $V$ and power law index $\beta$ (see Model in Material and Methods) by using a sliding window approach. The total number of time points was divided into subsets of five points, where the following subset was defined by adding the next time sampling and eliminating the earliest one. Both parameters were calculated for each subset against the average time lapse. Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the largest sampling: two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot). Both samples showed changes in the variability V with quasi--periodic behavior peaking at about 10 days. Variability grew more for the gut microbiota of the male and shared a minimal value of around 0.1 with the gut microbiota of the female.

>>>>

The microbiome temporal variability was analyzed to extract the global properties of the system. As fluctuations in total counts are plagued by systematic errors we worked on the temporal variability of relative abundances for each taxon. Our first finding was that, in all cases, changes in the relative abundances of taxa followed a ubiquitous pattern, known as the fluctuation scaling law\cite{fs} or Taylor's power law\cite{taylor}, i.e., the microbiota of all detected taxa followed $\sigma_i  = V\cdot x_i^{\beta}$, a power law dependence between the mean relative abundance $x_i$ and the dispersion $\sigma_i$. The law seem to be ubiquitous, spanning even to six orders of magnitude in the observed relative abundances. As shown in Figure \ref{fig:main1}, the most abundant species were less volatile in relative terms than those which were less abundant. The fitting to the power law was always robust ($R^{2}$ > 0.88) and did not depend on the microbiome condition. The power law (or scaling) index $\beta$ and the variability $V$ (hereafter Taylor's parameters) appear to be correlated with the stability of the community. On the one hand, $\beta$ is a scaling index that gave us information about the statistical properties of the ecosystem. If it is $1/2$, the system behaves like a Poisson distribution. If $\beta$ is 1, the system behaves as an exponential distribution. Generally speaking, metagenomes vary with time with $\beta$ between these two universal classes. In our case, the fact that $\beta$ was less than 1 tells us that the most abundant taxa in the microbial community were less susceptible to any perturbation than the other taxa. On the other hand, the variability $V$ was a direct estimator of the amplitude of fluctuations over time. $V$ represents the maximum variability attainable by a hypothetical dominant genus (with relative abundance close to 1). It is an important parameter that characterizes the type of system. If $V$ is small the ranking is stable such as, for example, the number of diagnoses of a particular disease recorded in Medicare during a month. If $V$ is large, as it seems to be the case of metagenomic samples, the ranking might be unstable, like the number of hourly page views of articles in Wikipedia \cite{ranking,fs}. The Taylor parameters were related to the health status of the host, which we consider as constituting the main finding contributed by this article.

\begin{figure}
	\centering
	\vspace*{-15mm} % Corrects overbox of the figures
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_h_A_amplicons_family_stdVSmean_xWboot_LOG.pdf}
	%\includegraphics[width=0.8\textwidth]{results/fits/IBS_P2_Metatranscriptores_stdVSmean_xWboot_LOG.pdf}
	\includegraphics[width=0.8\textwidth]{figs/Fig1.eps}
	\caption{X-weighted power-law fits of the standard deviations versus the mean values for each bacterial genus monitored over time. The fit is shown for samples from a healthy subject (top) and from a subject diagnosed with irritable bowel syndrome (bottom), studied in our lab \cite{IBS}. Taylor's power law seems to be ubiquitous, spanning to six orders of magnitude. The error bars (\emph{mean-axis}) are the SEM.}
	\label{fig:main1}
\end{figure}

Taylor's parameters describing the temporal variability of the gut microbiome in our sampled individuals are shown in Supplementary Tables S\ref{tab:diet} to S\ref{tab:LEA}. Our results hint at ubiquitous behavior. Firstly, the variability (which corresponds to the maximum amplitude of fluctuations) was large, which suggests the resilient capacity of the microbiota. Secondly, the scaling index was always smaller than one, which means that more abundant taxa were less volatile than less abundant ones. In addition, Taylor's parameters for the microbiome of healthy individuals in different studies were compatible with estimated errors. This enabled us to define an area in the Taylor parameter space that we called the \emph{healthy zone}.

In order to jointly visualize and compare the results of individuals from different studies\cite{IBS,moving,antibiotic,LEA,kwashiorkor,diet,hostlife}, their Taylor parameters were standardized, with standardization meaning that each parameter was subtracted by the mean value and divided by the standard deviation of the group of healthy individuals for every single study independently. Due to the different systematics in each study, we defined a healthy region for each one of them, standardized to mean zero and variance one and computed mean and variance of ?unhealthy? with this standardization (for details of the procedure, please see Standardization subsection in Material and Methods). Therefore, different studies were isolated so that individuals from a given study did not affect the results on the ?unhealthy? individuals of the other studies. We think this statistical approach was safer, as we avoided to combine data with very different systematic errors. The healthy zone and the standardized Taylor parameters for individuals whose gut microbiota was compromised (i.e., they were suffering from IBS, kwashiorkor, altered diet, intake of antibiotics, a Salmonella infection, or had gone on a trip abroad) are shown in Figure \ref{fig:main2}. The variability in children developing kwashiorkor was smaller than that of their healthy twins. A meat/fish-based diet significantly increased variability when compared to a plant-based diet. All other cases presented increased variability, which was particularly severe and statistically significant at over 95\% confidence level (CL), for grade III obese patients on a diet, individuals taking antibiotics, the subject who had a Salmonella infection, the subject who did a travel abroad or the IBS-diagnosed patients. One global property emerged from all the worldwide data collected: the Taylor's parameters characterized the statistical behavior of microbiome changes. Furthermore, we verified that our conclusions were robust to systematic errors as a result of taxonomic assignment (see Taxa level selection in Material and Methods).

\begin{figure}
	\centering
	%\includegraphics[width=0.9\textwidth]{results/finalplot11.pdf}
	\includegraphics[width=1.0\textwidth]{figs/Fig2.eps}
	\caption{Taylor's law parameter space. All the data studied in this work were compiled here. The colored circle corresponds to a 68\% confidence level (CL) region of healthy individuals in the Taylor's parameter space, while the dashed line delimites the 98\% CL region. Points with errors place gut microbiome in the Taylor's parameter space, for each individual whose microbiota was compromised. It should be noted that the parameters were standardized (standard deviation units) to the healthy group in each study for every single study independently, for demonstrative and comparative purposes.}
	\label{fig:main2}
\end{figure}

Taylor's power law has been explained in terms of various effects, though none have brought a general consensus. It can be shown to have its origin in mathematical convergence which is similar to the central limit theorem, and thus virtually any statistical model designed to produce a Taylor law converges to a Tweedie distribution\cite{stat}, providing a mechanistic explanation based on the statistical theory of errors\cite{convergence1,convergence2,convergence3}. To reveal the generic mechanisms that drive different scenarios in the $\beta-V$ space, we modeled the system by assuming that taxon relative abundance followed a Langevin equation with, on the one hand, a deterministic term that captured the fitness of each taxon and, on the other hand, a randomness term associated with Gaussian random noise\cite{ranking}.Both terms were modeled by power laws, with coefficients that can be interpreted as the taxon fitness $F_i$ and the variability $V$ (see Model in Material and Methods). Fitness $F_i$ captures the time scale that the system needs to reach equilibrium (the size of variability $V$ may or may not allow to reach it). $F_i$ has dimensions of 1/time and roughly corresponds to the half-life of the system when decaying to the stable state. In fact, it is exactly the half-life if $\beta$ is one and $V$ is negligible. In this model, when $V$ is sufficiently low, abundances are stable in time. Differences in the variability $V$ can induce a noise-induced phase transition in the relative abundances of taxa. The temporal evolution of the probability of a taxon having the abundance $x_i$, given its fitness, is governed by the Fokker--Planck equation. The results of solving this equation show that stability is best captured by a phase space determined by the fitness $F$ and the amplitude of fluctuations $V$ (see Figure \ref{fig:main3}).

@@

The model predicted two phases for the gut microbiome: a stable phase with large variability that enabled some changes in the relative abundances of taxa; and an unstable phase with larger variability, above the phase transition, where the order of abundant taxa varies significantly over time. The microbiome of all healthy individuals was found to be in the stable phase, while the microbiome of several other individuals was shown to be in the unstable phase. In particular, individuals taking antibiotics and the IBS--diagnosed patient \emph{P2} had the most severe symptoms. In this phase diagram, each microbiota state is represented by a point at its measured variability $V$ and inferred fitness $F$. The model predicted high average fitness for all taxa, i.e., taxa were narrowly distributed in $F$. The fitness parameter was chosen with different values for demonstrative purposes. Fitness was larger for the healthiest subjects and smaller for the IBS--diagnosed patients.

@@

We also found those \emph{rank stability islands} for medium-ranked taxa in the other periods belonging to the individual \emph{A} in the host lifestyle study \cite{hostlife} (see Supplementary Figure S\ref{supfig:corrank_HLS_before} and Supplementary Figure S\ref{supfig:corrank_HLS_after} for the corresponding rank plots). See Supplementary Table S\ref{tab:HLS_RSI} for details about the rank and RSI for the above-mentioned taxa over the different periods considered.

@@

Finally, we studied the time dependence of the variability $V$ and power law index $\beta$ (see Model in Material and Methods) by using a sliding window approach. The total number of time points was divided into subsets of five points, where the following subset was defined by adding the next time sampling and eliminating the earliest one. Both parameters were calculated for each subset against the average time lapse. Figure \ref{fig:tempevo1} shows the variability $V$ as a function of time for the largest sampling: two individuals in Caporaso's study\cite{moving} corresponding to the gut microbiota of a male (upper plot) and a female (lower plot). Both samples showed changes in the variability $V$ with quasi--periodic behavior peaking at about 10 days. Variability grew more for the gut microbiota of the male and shared a minimal value of around 0.1 with the gut microbiota of the female.
